### 2025-09-22 Neutrino.tex
upgrade cls
neutirno and comosologgasdafafasd

\paragraph{Connection of the above two models}
If we set $M_N =0$ in the Type-I see-saw ($n\w{RHN}=3$), the AT-diagonalization becomes
\begin{align}
\tilde M = \frac12\pmat{0 & (U_nM\w{D}\wdiag V_n^\dagger)^\TT \\ U_nM\w{D}\wdiag V_n^\dagger & 0},\quad
  \tilde R=\pmat{V_n^* & \ii V_n^* \\ U_n & -\ii U_n},\quad
  \tilde M\wdiag=\pmat{M\w{D}\wdiag & 0 \\ 0 & M\w{D}\wdiag}.
\end{align}
As three pairs of degenerate Weyl fermions form three Dirac neutrinos, the Dirac neutrino model is reproduced as expected.

@@

\subsection{PMNS matrix}\label{sec:PMNS}
As experiments have not found any deviation from $[U]\w{upper left}$ being unitary, we express the mixing with the PMNS matrix
 \GRAY{(Pontecorvo--\JAPANESE{牧}--\JAPANESE{中川}--\JAPANESE{坂田})}
parameterized by
\begin{equation}
\begin{split}
   U\w{PMNS}&=
 \pmat{1&&\\&\co{23}&\si{23}\\&-\si{23}&\co{23}}
 \pmat{\co{13}&&\si{13}\ee^{-\ii\delta\w{CP}}\\&1&\\-\si{13}\ee^{\ii\delta\w{CP}}&&\co{13}}
 \pmat{\co{12}&\si{12}&\\-\si{12}&\co{12}&\\&&1}
\pmat{\ee^{\ii\eta_1} \\ & \ee^{\ii\eta_2}\\&&1}
 %
\\&=
 \pmat{
 \co{12} \co{13} & \si{12} \co{13} & \si{13} \ee^{-\ii\delta\w{CP}}\\
 -\si{12} \co{23} - \co{12} \si{23} \si{13} \ee^{\ii\delta\w{CP}}& \co{12} \co{23} - \si{12}\si{23}\si{13} \ee^{\ii\delta\w{CP}}& \si{23}\co{13}\\
  \si{12}\si{23} - \co{12} \co{23} \si{13}\ee^{\ii\delta\w{CP}} & -\co{12}\si{23}-\si{12}\co{23}\si{13}\ee^{\ii\delta\w{CP}} & \co{23} \co{13}
}\pmat{\ee^{\ii\eta_1} \\ & \ee^{\ii\eta_2}\\&&1},
\end{split}
\end{equation}
where $\theta_{ij}\in[0,\pi/2]$ and $\delta\w{CP}\in[0,2\pi]$ as shown in \cref{eq:GeneralUnitary33}.
In the Dirac neutrino model, $\eta_1$ and $\eta_2$ (``Majorana phases'') are unphysical because we can rotate $\nR$ to eliminate them, while the rotation is not allowed in presence of the Majorana mass term.

>>>>

\paragraph{Connection of the above two models}
If we set $M_N =0$ in the Type-I see-saw ($n\w{RHN}=3$), AT-diagonalization results in 
\begin{align}
\tilde M = \frac12\pmat{0 & (U_nM\w{D}\wdiag V_n^\dagger)^\TT \\ U_nM\w{D}\wdiag V_n^\dagger & 0},\quad
  \tilde R=\pmat{V_n^* & \ii V_n^* \\ U_n & -\ii U_n},\quad
  \tilde M\wdiag=\pmat{M\w{D}\wdiag & 0 \\ 0 & M\w{D}\wdiag}
\end{align}
and we have three pairs of degenerate Weyl fermions, i.e., three Dirac neutrinos, which is equivalent to the Dirac neutrino model.

\subsection{Weinberg Operator}
The neutrino masses are often described by the Weinberg operator~\cite{Weinberg:1979sa}, which is obtained by integrating out the right-handed neutrinos. It is given by, usually with the convention $C_{ij}=C_{ji}$ because of $(L_i H)(L_j H)=(L_j H)(L_i H)$,
\begin{equation}
  \mathcal L\ni \frac{C_{ij}}{2}(L_i H)(L_j H)+\text{H.c.}
  \equiv
  \frac{C_{ij}}{2}\epsilon^{ab}\epsilon^{cd}(L_i^a H^b)^\alpha(L_j^c H^d)_\alpha+\text{H.c.}
  \leadsto
  \frac{C_{ij}}{2}\mean{H^0}^2(\nu_i\nu_j)+\text{H.c.}
\end{equation}
The (Majorana) mass matrix is thus given by $C_{ij}\mean{H^0}^2=C_{ij}v^2/2$ with $v=246\GeV$.
The overall sign is unphysical here but may appear in the PMNS matrix.

@@

\subsection{PMNS matrix}\label{sec:PMNS}
As experiments have not found any deviation from $[U]\w{upper left}$ being unitary, we express the mixing with the PMNS matrix
 \GRAY{(Pontecorvo--\JA{牧}--\JA{中川}--\JA{坂田})}
parameterized by
\begin{equation}
\begin{split}
   U\w{PMNS}&=
 \pmat{1&&\\&\co{23}&\si{23}\\&-\si{23}&\co{23}}
 \pmat{\co{13}&&\si{13}\ee^{-\ii\delta\w{CP}}\\&1&\\-\si{13}\ee^{\ii\delta\w{CP}}&&\co{13}}
 \pmat{\co{12}&\si{12}&\\-\si{12}&\co{12}&\\&&1}
\pmat{\ee^{\ii\eta_1} \\ & \ee^{\ii\eta_2}\\&&1}
 %
\\&=
 \pmat{
 \co{12} \co{13} & \si{12} \co{13} & \si{13} \ee^{-\ii\delta\w{CP}}\\
 -\si{12} \co{23} - \co{12} \si{23} \si{13} \ee^{\ii\delta\w{CP}}& \co{12} \co{23} - \si{12}\si{23}\si{13} \ee^{\ii\delta\w{CP}}& \si{23}\co{13}\\
  \si{12}\si{23} - \co{12} \co{23} \si{13}\ee^{\ii\delta\w{CP}} & -\co{12}\si{23}-\si{12}\co{23}\si{13}\ee^{\ii\delta\w{CP}} & \co{23} \co{13}
}\pmat{\ee^{\ii\eta_1} \\ & \ee^{\ii\eta_2}\\&&1},
\end{split}
\end{equation}
where $\theta_{ij}\in[0,\pi/2]$ and $\delta\w{CP}\in[0,2\pi]$ as shown in \cref{eq:GeneralUnitary33}.
In the Dirac neutrino model, $\eta_1$ and $\eta_2$ (``Majorana phases'') are unphysical because we can rotate $\nR$ to eliminate them, while the rotation is not allowed in presence of the Majorana mass term.
